---
title: "R Notebook"
---

1. Load in the tidyverse package 
```{r}
library(tidyverse)
```

2. Load and view the mtcars data

```{r}
cars <- mtcars
cars$model <- rownames(cars)

cars
```

Variable descriptions: 
mpg: Miles/(US) gallon
cyl: Number of cylinders
disp: Displacement (cu.in.)
hp: Gross horsepower
drat: Rear axle ratio
wt: Weight (1000 lbs)
qsec: 1/4 mile time
vs: V/S
am: Transmission (0 = automatic, 1 = manual)
gear: Number of forward gears
carb: Number of carburetors

3. Plot weight (wt) against miles per gallon (mpg)
```{r}
ggplot(data = cars,
       mapping = aes(x = wt,
                     y = mpg)) + 
  geom_point()
```

4. Colour the dots according to displacement (disp) value
```{r}
ggplot(data = cars,
       mapping = aes(x = wt,
                     y = mpg,
                     colour = disp)) + 
  geom_point()
```

What is the car with the highest weight and miles per gallon? Let's try to add some labels. 

Note:
Most of the time, all the data for a given plot are contained in one table. This means that the mapping can be done in the main plotting call `ggplot()` and does not have to go into layers. 

```{r}
ggplot(data = cars,
       mapping = aes(x = wt,
                     y = mpg,
                     colour = disp)) + 
  geom_point() + 
  geom_text(mapping = aes(label = model))
```

The labels are probably overlapping, how can we see what's the highest? 
```{r}
ggplot(data = cars,
       mapping = aes(x = wt,
                     y = mpg,
                     colour = disp)) + 
  geom_point() + 
  geom_text(data = cars %>% filter(wt == max(wt)),
            mapping = aes(label = model),
            nudge_y = 1,
            nudge_x = -0.3)
```

Add labels and title to the plot

```{r}
ggplot(data = cars,
       mapping = aes(x = wt,
                     y = mpg,
                     colour = disp)) + 
  geom_point() + 
  geom_text(data = cars %>% filter(wt == max(wt)),
            mapping = aes(label = model),
            nudge_y = 1,
            nudge_x = -0.3) + 
  labs(x = "weight",
       y = "miles per gallon",
       title = "Weight vs. miles per gallon in mtcars dataset")
```




