---
title: "R Notebook"
---

Tutorial credits: https://www.biostars.org/p/458364/

```{r}
library(tidyverse)
```

Let's create a sample table. Many biological results can be in this format:
```{r}
mat <- cbind(
  t(replicate(4, rnorm(3, runif(1, 10, 500), runif(1, 1, 25)))),
  t(replicate(4, rnorm(3, runif(1, 10, 500), runif(1, 1, 25)))))

dimnames(mat) <- list(
  sprintf("ENS%06d", seq(100, 103)),
  c(sprintf("EV_%02d", seq_len(3)), sprintf("KD_%02d", seq_len(3))))

df <- as_tibble(mat, rownames="genes")
df
```

Unfortunately, ggplot does not handle data in "wide" format well. It is simply impossible to map values to plot components. We have to transform it into a "long" format, where every row is a separate observation. 

```{r}
df <- pivot_longer(df, -genes, names_to = "sample", values_to = "expression")
df
```

This looks better! But we still have multiple replicates per sample, so we could separate these into two columns. 

```{r}
df <- separate(df, sample, into=c("sample", "replicate"), sep="_")
df
```

Amazing, ggplot loves this! Now we are ready to plot the data.

```{r}
ggplot(df, 
       aes(x = genes, 
           y = log2(expression), 
           fill = sample)) +
  geom_boxplot()
```

Sometimes the bottom labels may be overlapping. This can be fixed in two ways: 
1. Rotating the labels 
2. Rotating the entire table

```{r}
ggplot(df, 
       aes(x = genes, 
           y = log2(expression), 
           fill = sample)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(df, 
       aes(x = genes, 
           y = log2(expression), 
           fill = sample)) +
  geom_boxplot() +
  coord_flip()
```
